---
title: "Job_applications"
authors: "Sajan Kumar Kar, Sai Pavan Mekala, Neelima Puthanveetil, Sandhya Karki"
date: "2023-10-15"
output:
  html_document:
    code_folding: hide
    theme: flatly
    highlight: pygments
    number_sections: false
    toc: yes
    toc_depth: 3
    toc_float: yes
  pdf_document:
    toc: yes
    toc_depth: '3'
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Team Global project on Job Applicants Data
## 1. Import Libraries


```{r Libraries}
# uncomment and run the below line to install the required packages
#install.packages(c("gganimate", "png", "gifski"))
library(ggplot2)
library(tidyr)
library(dplyr)
library(ezids)
library(gganimate)
library(png)
library(gifski)

```

## 2. Import Data

```{r Import_data}
df = data.frame(read.csv('stackoverflow_full.csv', header = TRUE))
xkabledplyhead(df, title = "Job Applicants Data")
```
```{r Structure}
#structure of a data frame 
str(df) 
```

The dataframe has 73462 records. We have `index`, `employemnt`, `yearscode`, `yearscodepro` and `previoussalary` as integer type and the rest are of character types. Going ahead we would need to convert some of them to factors.

## 3. Data preprocessing & Cleaning

Let's first check for missing values and duplicate values if any.

```{r missing values}
# Check the missing values and count them in each column
print(colSums(is.na(df)))
```


```{r duplicates}
# Check duplicates
length(unique(df$Index))
```

We notice that there are no missing values and duplicates in the data. So, we have all the observations to work with.

Now that we know this, let us do some cleaning and feature engineering. First we'll extract the numeric columns.

```{r Numeric columns}
# Extract the numeric columns
head(subset(df, select = c(names(df)[sapply(df, is.numeric)])))
```


```{r numeric feature cleaning}
df_clean <- subset(df, select = -c(Index))
df_clean$Employment = as.character(df_clean$Employment)
df_clean$Employment[df_clean$Employment == "1"] <- "currently_employed"
df_clean$Employment[df_clean$Employment=="0"] <- "not_currently_employed"
df_clean$Employment = factor(df_clean$Employment)
df_clean$Employed = as.character(df_clean$Employed)
df_clean$Employed[df_clean$Employed == "1"] <- "hired"
df_clean$Employed[df_clean$Employed == "0"] <- "not_hired"
df_clean$Employed = factor(df_clean$Employed)
str(df_clean)
```

Now let's take a look at the character columns.

```{r char columns}
# Extract the character columns
head(subset(df_clean, select = c(names(df)[sapply(df, is.character)])))

```

```{r char feature cleaning}
df_clean$Age = factor(df_clean$Age)
df_clean$Accessibility = factor(df_clean$Accessibility)
df_clean$EdLevel = factor(df_clean$EdLevel, ordered = T, 
                          levels = c("Other", "NoHigherEd", "Undergraduate", "Master", "PhD"))
df_clean$Gender = factor(df_clean$Gender)
df_clean$MentalHealth = factor(df_clean$MentalHealth)
df_clean$MainBranch = factor(df_clean$MainBranch)
str(df_clean)
```


Let's check the number of different countries we have.

```{r country}
length(unique(df_clean$Country))

```

There are 172 different countries. (Leaving this as it is for the time being)

Let's see if there is data where their professional coding experience is more than their coding experience. If there is such data, we have to remove it because that's practically faulty data.

```{r remove faulty data}

nrow(subset(df_clean, subset = df_clean$YearsCode<df_clean$YearsCodePro))
# We observe 588 such records
df_clean = subset(df_clean, subset = df_clean$YearsCode>df_clean$YearsCodePro)

```


Let's take a look at some plots and distribution in order to deal with outliers.

```{r outlier detection}
# Create a list of the columns that we use in combined boxplot
columns_to_plot <- c("YearsCode", "YearsCodePro", "PreviousSalary", "ComputerSkills")

# Set up the plotting area to display all boxplots together
par(mfrow = c(2, length(columns_to_plot)))

# Display the boxplots
for (i in 1:length(columns_to_plot)) {
  boxplot(df_clean[[columns_to_plot[i]]], main = columns_to_plot[i],col = "lightblue")
}
for (i in 1:length(columns_to_plot)) {
  hist(df_clean[[columns_to_plot[i]]], main = columns_to_plot[i], xlab = element_blank(), col = "lightblue")
}
# plotting area layout
par(mfrow = c(2, 1))

```

```{r quantiles}
o_code= quantile(df_clean$YearsCode, probs = 0.75, na.rm = FALSE) + 
  (1.5 * IQR(df_clean$YearsCode))
o_procode = quantile(df_clean$YearsCodePro, probs = 0.75, na.rm = FALSE) + 
  (1.5 * IQR(df_clean$YearsCodePro))
o_presal = quantile(df_clean$PreviousSalary, probs = 0.75, na.rm = FALSE) + 
  (1.5 * IQR(df_clean$PreviousSalary))
o_cskill = quantile(df_clean$ComputerSkills, probs = 0.75, na.rm = FALSE) + 
  (1.5 * IQR(df_clean$ComputerSkills))

print(paste(o_code, o_procode, o_presal, o_cskill))
```

```{r outlier removal}
df_clean = subset(df_clean, subset = df_clean$YearsCode < 40)
df_clean = subset(df_clean, subset = df_clean$YearsCodePro < 40)
df_clean = subset(df_clean, subset = df_clean$ComputerSkills < 35)
```


Lets take a look at the final data now:

```{r final data}
df_final <- df_clean
xkabledplyhead(df_final, title = "Final Data")

```

```{r Final Data Basic Info}

# Getting the number of categorical and numerical columns
cat_cols <- colnames(subset(df_final, select = c(names(df_final)[(sapply(df_final, is.factor)) | 
                                                       (sapply(df_final, is.ordered)) | 
                                                       (sapply(df_final, is.character))])))

num_cols <- colnames(subset(df_final, select = c(names(df_final)[sapply(df_final, is.integer)])))

# Displaying the information
cat("Basic summary and Raw Counts for the Dataset:\n")
cat("Rows:", dim(df_final)[1], "\n")
cat("Columns:", dim(df_final)[2], "\n")
cat("Numerical columns:", num_cols, "\n")
cat("Categorical Columns:", cat_cols, "\n")

```

## 4. EDA

Let's first try to get a good picture of the data which will act as the 1st step to answer the questions that we proposed.


```{r plot- Ed by Gender}

ggplot(data = subset(df_final, subset = Employed=='hired'), aes(EdLevel, after_stat(count))) + 
  geom_bar(aes(fill = Gender), position = 'dodge', alpha = 0.5) +
  labs(title = "Education Level by gender for employed workers", x="Education Level", y="Count")

```


```{r plot- Ed by Emp}
ggplot(data = df_final, aes(EdLevel)) + 
  geom_bar(aes(fill = Employment), position = 'dodge', alpha = 0.5) +
  labs(title = "Education Level by Employment", x="Education Level", y="Count")

```


```{r plot- MainBranch by Age}
ggplot(data = df_final, aes(MainBranch)) + 
  geom_bar(aes(fill = Age), position = 'dodge', alpha = 0.5) +
  labs(title = "MainBranch by Age", x="MainBranch", y="Count")

```


```{r plot- PrevSal by MentalHealth(1)}
ggplot(data = df_final, aes(x = PreviousSalary, fill = MentalHealth)) +
  geom_density(alpha = 0.5) +
  labs(title = "Density plot of Previous Salary by Mental Health", x="Previous Salary", y= 'Density')

```


```{r plot- PrevSal by MentalHealth(2)}

ggplot(data = df_final, aes(x = MentalHealth, y = PreviousSalary)) +
  geom_jitter(alpha = 0.5, col = 'navy') +
  labs(title = "Previous Salary by Mental Health", x = 'Mental Health', y="Previous Salary")

```

```{r plot- PrevSal by MentalHealth(3)}
plotdata <- df_final %>%
  group_by(MentalHealth) %>%
  summarize(mean_prevsalary = mean(PreviousSalary))

ggplot(data = plotdata, aes(x = MentalHealth, y = mean_prevsalary)) +
  geom_bar(stat = 'identity', alpha = 0.5, aes(fill = MentalHealth)) +
  labs(title = "Mean Previous Salary by Mental Health", x = 'Mental Health', y="Previous Salary")

```
We notice that the mean previous salary of applicants having mental health issues is higher.


Let's try digging a bit deeper into the data to get additional insights.
Let's try to extract the most used skill-
```{r skills extraction}

df_final %>%
separate_rows(HaveWorkedWith, sep = ";") %>% 
group_by(HaveWorkedWith) %>%
summarise(Count = n()) %>%
arrange(desc(Count)) -> skills_df
head(skills_df, 10)
```

We see that Javascript, Docker and HTML/CSS are the skills known by majority of the applicants. Let's visualize it to have a better look.

```{r plot- Top 10 skills known}

plot <- ggplot(data = head(skills_df, 10), aes(x = reorder(HaveWorkedWith, Count), y = Count)) + 
  geom_bar(stat = 'identity', aes(fill = HaveWorkedWith), alpha = 0.6) + 
  coord_flip() + 
  theme(axis.text.x = element_blank(),
        axis.line = element_blank(),
        axis.ticks = element_blank(),
        panel.background = element_blank(),
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        panel.grid.major.x = element_line(size = 0.1, color = "grey"),
        panel.grid.minor.x = element_line(size = 0.1, color = "grey"),
        plot.background = element_blank()) + 
  geom_text(aes(y = Count, label = Count)) +
  labs(title = "Skills used by number of applicants", 
       x = "Skill", y = "No. of applicants possessing the skill")

anim <- plot + 
    transition_states(HaveWorkedWith, transition_length = 4, wrap = FALSE) + 
  shadow_mark() + 
  enter_grow() +
  enter_fade() + 
  ease_aes('sine-in')
  
animate(anim)
#anim_save("Skills.gif", anim)
```
Because we have over 65,000 data points, we can use statistical methods that apply to data satisfying normality conditions (because of the central limit theorem).


## 5. Hypothesis Testing
We consistently use the standard alpha-level of .05 as a basis for making conclusions for our hypothesis tests.

### Test 1: Is there a significant difference between employed males and non-males in their respective education levels?

$H_0:$ There is no difference in the number of undergraduate males who are employed when compared to nonmales

$H_a:$ There is a significant difference in the number of undergraduate males who are employed when compared to nonmales
```{r, two sample z test for proportions: undergraduate}
# We first subset into the different education levels, here undergraduate
undergrad <- subset(df_clean, df_clean$EdLevel == "Undergraduate")


# For the two sample z test, we need our two samples, males and nonmales
undergrad_male <- subset(undergrad, undergrad$Gender == "Man") # Sample 1, males

undergrad_male_employed <- subset(undergrad_male, undergrad_male$Employment == "currently_employed")

undergrad_nonmales <- subset(undergrad, undergrad$Gender != "Man") # Sample 2, females and nonbinary

undergrad_nonmales_employed <- (subset(undergrad_nonmales, undergrad_nonmales$Employment == "currently_employed"))


undergrad_employed_ztest <- prop.test(c(nrow(undergrad_male_employed), nrow(undergrad_nonmales_employed)), c(nrow(undergrad_male),nrow(undergrad_nonmales)))
undergrad_employed_ztest
```
At an alpha-level of .05 we fail to reject the null hypothesis and conclude that there is not a significant difference in the number of males vs nonmales who have a bachelor's degree and are employed.

$H_0:$ There is no difference in the number of males with a master's degree who are employed when compared to nonmales

$H_a:$ There is a significant difference in the number of males with a master's degree who are employed when compared to nonmales

```{r, two sample z test for proportions: masters}
# We first subset into the different education levels, here masters
masters <- subset(df_clean, df_clean$EdLevel == "Master")


# For the two sample z test, we need our two samples, males and nonmales
masters_male <- subset(masters, masters$Gender == "Man") # Sample 1, males

masters_male_employed <- subset(masters_male, masters_male$Employment == "currently_employed")


masters_nonmales <- subset(masters, undergrad$Gender != "Man") # Sample 2, females and nonbinary

masters_nonmales_employed <- (subset(masters_nonmales, masters_nonmales$Employment == "currently_employed"))


masters_employed_ztest <- prop.test(c(nrow(masters_male_employed), nrow(masters_nonmales_employed)), c(nrow(masters_male),nrow(masters_nonmales)))
masters_employed_ztest

```
At and alpha-level of .05 we reject the null hypothesis and conclude that there is a significant difference in the number of males vs nonmales who have a master's degree and are employed.

$H_0:$ There is no difference in the number of males with a PhD who are employed when compared to nonmales

$H_a:$ There is a significant difference in the number of males with a PhDwho are employed when compared to nonmales
```{r, two sample z test for proportions: Phd}
# We first subset into the different education levels, here PhD
phd <- subset(df_clean, df_clean$EdLevel == "PhD")


# For the two sample z test, we need our two samples, males and nonmales
phd_male <- subset(phd, phd$Gender == "Man") # Sample 1, males

phd_male_employed <- subset(phd_male, phd_male$Employment == "currently_employed")

phd_nonmales <- subset(phd, phd$Gender != "Man") # Sample 2, females and nonbinary

phd_nonmales_employed <- (subset(phd_nonmales, phd_nonmales$Employment == "currently_employed"))


phd_employed_ztest <- prop.test(c(nrow(phd_male_employed), nrow(phd_nonmales_employed)), c(nrow(phd_male),nrow(phd_nonmales)))
phd_employed_ztest
```
At and alpha-level of .05 we fail to reject the null hypothesis and conclude that there is not a significant difference in the number of males vs nonmales who have a doctorate degree and are employed.

### Test 2: Does education level significantly impact employment in the technical field?
```{r, ANOVA boxplots}
loadPkg("ggplot2")
ggplot(df_clean, aes(x=EdLevel, y=ComputerSkills)) + 
  geom_boxplot( colour=c("#ff0000","#FFFF00", "#11cc11","#0000ff","#ff00ff"), alpha = .5, outlier.shape=8, outlier.size=4) +
  labs(x="Education Level", y = "Number of Computer Skills")
```

$H_0:$: There is no significant difference in the average number of computer skills acquired when comparing all individuals at every level of education.

$H_a:$ At least one education level group differs from the others in the average number of computer skills known.

```{r, ANOVA test}
anovaRes = aov(ComputerSkills ~ EdLevel, data=df_clean)
anovaSummary = summary(anovaRes)
anovaSummary

# Extracting mean computer skills for each education level
mean_skills_Other = mean(subset(df_clean, EdLevel == "Other")$ComputerSkills)
mean_skills_NoHigherEd = mean(subset(df_clean, EdLevel == "NoHigherEd")$ComputerSkills)
mean_skills_Undergraduate = mean(subset(df_clean, EdLevel == "Undergraduate")$ComputerSkills)
mean_skills_Master = mean(subset(df_clean, EdLevel == "Master")$ComputerSkills)
mean_skills_PhD = mean(subset(df_clean, EdLevel == "PhD")$ComputerSkills)
```
At an alpha-level of .05, we reject the null hypothesis. We conclude that there is at least one education level for which the average number of computer skills is different from the other groups.

```{r, Post Hoc Tukeys Test}
tukeyRes <- TukeyHSD(anovaRes)
tukeyRes
```

Since all p-values for every pairwise comparison was less than 0.05, we conclude that the average number of computer skills for every education level was significantly different from every other level of education.

### Test 3: Is age a significant factor in determining whether an individual is a professional developer or not?

$H_0:$ Whether an individual is a professional developer or not is independent of their age.

$H_a:$ Age and status of being a professional developer are associated with each other.
```{r, Chi square}
# We need four categories: <35, NotDev; >35, NotDev; <35 Dev; >35 Dev
notDev_under35 <- subset(df_clean, df_clean$Age == "<35" & df_clean$MainBranch == "NotDev")
notDev_over35 <- subset(df_clean, df_clean$Age == ">35" & df_clean$MainBranch == "NotDev")
dev_under35 <- subset(df_clean, df_clean$Age == "<35" & df_clean$MainBranch == "Dev")
dev_over35 <- subset(df_clean, df_clean$Age == ">35" & df_clean$MainBranch == "Dev")
```

The contingency table that we perform our chi square test on is the following:
```{r, Chi square table}
chisq_matrix <- matrix(c(nrow(dev_over35),nrow(dev_under35),nrow(notDev_over35),nrow(notDev_under35)), nrow = 2, ncol = 2)

# Now we conduct a chi square test of independence
chitest = chisq.test(chisq_matrix)
chitest
```
At an alpha-level of .05 we reject the null hypothesis. We conclude that age and being a professional developer are associated.

### Test 4: Does mental health influence previous salary?
$H_0:$ Whether an individual is a professional developer or not is independent of their age.

$H_a:$ Age and status of being a professional developer are associated with each other.
```{r, two sample t-test for means}
# Subset the PreviousSalary for individuals with 'Yes' and 'No' in the MentalHealth column
salary_yes <- df_clean$PreviousSalary[df_clean$MentalHealth == "Yes"]
salary_no <- df_clean$PreviousSalary[df_clean$MentalHealth == "No"]

# Checking the mean of both groups just to get an idea
format(mean(salary_yes, na.rm = TRUE), digits=5)
format(mean(salary_no, na.rm = TRUE), digits=5)

# Now we conduct a two-sample t-test on PreviousSalary between the two groups
ttest_2sample_salary_yes_no <- t.test(salary_yes, salary_no)
ttest_2sample_salary_yes_no

```
At an alpha-level of .05 we reject the null hypothesis. We conclude that the mean salary of those with mental health issues is significantly different (in this case, higher) than those without.

## 6. Data Preparation


## 7. Model Building

